Continuous Deductive Reasoning over Streamed RDF Data

### 2015-07-08 content/contribution/continuous-datalog.tex
Added a summary of the relationship between Continuous Datalog programs, their states and conventional Datalog programs, and how the same relationships apply between their entailments.

The entailment of such a program is composed of a stream of entailed
instances $P$ and a stream of instance negations $N$, whose contents
are defined in 
Axioms~\ref{axiom:continuous datalog: positive window increment}, 
\ref{axiom:continuous datalog: negative window increment} and
\ref{axiom:continuous datalog: entailment precedes negation}. 
These specify that the streams $P$ and $N$ are the continuous
reasoning equivalents of the streams produced by the \emph{IStream}
and \emph{DStream} operators of the \emph{CQL} continuous query
language \citep{CQL}, respectively.
\[ \Pi^C = \langle R, F, S, r \rangle \vDash \langle P, N \rangle \]
The truth value of a $\Pi^C$ is instantaneous, with the the truth
function of Continuous Datalog $T^{CD}$ taking both a program ${\Pi^C}$
and an instant ${i \in \mathbb{N}}$ as arguments.
\end{definition}

@@

Thus far the translation from instantaneous sets of entailments to
streams of entailments has been stated axiomatically as the definition
of those streams.  However, it can be shown from these axioms, as in
\ref{sec:semantics: proof for entailments}, that it is also possible
to reconstruct exactly the instantaneous set $E^T_{i}$, of instances
entailed at any given instant $i$ by any $\Pi^C = \{ R, F, S, r \}$,
from the sliding windows of sufficient range over its entailed streams
$P$ and $N$ at $i$:
\[ E^T_{i} = W^{P,x}_{i} \setminus W^{N,y}_{i} \]
where ${y \geq x \geq r}$. The minimum range $x$ is such to achieve
the \emph{complete} set of entailment instances $E^T_{i}$ of the
$\Pi^C_{i}$. In addition, the minimum range $y$ is such to achieve the
\emph{correct} set $E^T_{i}$. As such, a \emph{valid window} that
contains only and exactly those entailed instances $E^T_{i}$ at every
instant $i$ may be trivially maintained from the streams $P$ and $N$
continuously entailed by some $\Pi^C$.
\end{nestedsection}

>>>>

The entailment of such a program is composed of a stream of entailed
instances $P$ and a stream of instance negations $N$, whose contents
are defined in 
Axioms~\ref{axiom:continuous datalog: positive window increment}, 
\ref{axiom:continuous datalog: negative window increment} and
\ref{axiom:continuous datalog: entailment precedes negation}. 
These specify that the streams $P$ and $N$ are the continuous
reasoning equivalents of the streams produced by the \emph{IStream}
and \emph{DStream} operators of the \emph{CQL} continuous query
language \citep{CQL}, respectively.
\[ \Pi^C = \langle R, F, S, r \rangle \vDash \langle E, P, N \rangle \]
The truth value of a $\Pi^C$ is instantaneous, with the the truth
function of Continuous Datalog $T^{CD}$ taking both a program ${\Pi^C}$
and an instant ${i \in \mathbb{N}}$ as arguments.
\end{definition}

@@

Thus far the translation from instantaneous sets of entailments to
streams of entailments has been stated axiomatically as the definition
of those streams.  However, it can be shown from these axioms, as in
\ref{sec:semantics: proof for entailments}, that it is also possible
to reconstruct exactly the instantaneous set $E^T_{i}$, of instances
entailed at any given instant $i$ by any $\Pi^C = \{ R, F, S, r \}$,
from the sliding windows of sufficient range over its entailed streams
$P$ and $N$ at $i$:
\[ E^T_{i} = W^{P,x}_{i} \setminus W^{N,y}_{i} \]
where ${y \geq x \geq r}$. The minimum range $x$ is such to achieve
the \emph{complete} set of entailment instances $E^T_{i}$ of the
$\Pi^C_{i}$. In addition, the minimum range $y$ is such to achieve the
\emph{correct} set $E^T_{i}$. As such, a \emph{valid window} that
contains only and exactly those entailed instances $E^T_{i}$ at every
instant $i$ may be trivially maintained from the streams $P$ and $N$
continuously entailed by some $\Pi^C$.

To sum up, a time-annotated Datalog program
${\Pi^C_i = \langle R, F^0 \cup W^{S,r}_i \rangle}$ may be derived from
a continuous Datalog program ${\Pi^C = \langle R, F, S, r \rangle}$ by
applying a sliding window $W^{S,r}_i$ of range $r$ over the stream $S$ at
instant $i$ and instantiating the facts in $F$ as $F^0$. At the same time,
the entailments of a ${\Pi^C_i \vDash E^0 \cup E^T_i}$ may be derived from
the entailments of the ${\Pi^C \vDash \langle E, P, N \rangle}$ as the union
of the entailments of ${\langle R, F^0 \rangle \vDash E^0}$ and the set
difference of the windows $W^{P,r}_i$ and $W^{N,r}_i$ of range $r$ over
streams $P$ and $N$ at instant $i$.
Furthermore, a conventional Datalog program ${\Pi = \langle R, F \rangle}$
may be derived from a time-annotated Datalog program $\Pi^C_i$ by
${strip}$ping the time annotations from the set of facts
${F^0 \cup W^{S,r}_i}$, as the entailments of ${\Pi \vDash E}$ may be
derived from the entailments of $\Pi^C_i$ by $strip$ping them.
This yields the TODO RMEMBER NAME OF GRAPH in \reffig{CD-TD-D}.
\begin{figure}
	\centering
	% TODO make the diagram
	\caption{TODO}
	\labelfig{CD-TD-D}
\end{figure}
\end{nestedsection}
